A Rough Guide to Pre-processing High-Frequency Animal Tracking Data

### 2020-10-12 05_calibration_data.Rmd
try recurse on boating data

# plot data
fig_data_raw <-
  ggplot(data)+
  geom_path(aes(x, y),
             col = "grey", alpha = 0.5, size = 0.3)+
  geom_point(aes(x, y),
             col = "steelblue", alpha = 0.2, size = 2)+
  ggthemes::theme_map()+
  coord_sf(crs = 32631)

@@

```{r}
# divide by 1000, convert to integer, then convert to POSIXct
data[, time := as.POSIXct(as.integer(TIME / 1000), 
                          origin = "1970-01-01")]
```

@@

```{r}
# plot filtered data
fig_speed_outliers <-
  ggplot()+
  geom_point(data = data,
             aes(x, y, 
                 col = speed_in < 24),
             size = 0.3, alpha = 0.5,
             show.legend = F)+
  scale_color_manual(values = c("grey", pal[3]))+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "grey99"))+
  ggspatial::annotation_scale(location = "tl")+
  coord_sf(crs = 32631)+
  labs(colour = "speed (m/s)")

# save
ggsave(fig_speed_outliers, filename = "figures/fig_speed_outlier.png",
       width = 185 / 25, height = 185 / 25)
```

@@

Here we use a speed threshold of 20 m/s, 4 m/s lower than the 95th percentile.

@@

# remove speed outliers
data <- atl_filter_covariates(data = data,
            filters = c("(speed_in < 15.7 & speed_out < 15.7)"))

@@

```{r}
# make a data copy
data_unproc <- copy(data)

# assign residence positions
data[, residence := stringi::stri_detect(tID, regex = "WP")]

# make a copy of only residence points
data_res <- data[residence == T, ]
```

@@

```{r}
# load recurse
library(recurse)

# get recurse data for a 10m radius
data_recurse <- getRecursions(data_unproc[, list(x, y, time, TAG)],
                              radius = 100)
```

@@

```{r}
# on known residence points
patch_res_known <- atl_res_patch(data_res, 
                                buffer_radius = 10,
                                lim_spat_indep = 100,
                                lim_time_indep = 30,
                                min_fixes = 3)
```

@@

```{r}
# patch with residence points and all patches
fig_basic_residence <-
  ggplot()+
  geom_path(data = data,
            aes(x, y),
            lwd = 0.2,
            col = "grey20")+
  geom_point(data = data_res,
             aes(x, y),
             shape = 2,
             size = 2,
             col = pal[2])+
  geom_path(data = patch_summary_data,
            aes(x_median + 200, y_median),
            col = "orange", lty = 1)+
  geom_point(data = patch_summary_data,
             aes(x_median + 200, y_median,
                 fill = duration / 60),
             size = 5,
             shape = 21)+

  scale_fill_distiller(palette = "YlOrRd",
                       direction = 1,
                       trans = "log10",
                       limits = c(1, 60))+
  ggthemes::theme_map()+
  theme(legend.position = c(0, 0.3),
        legend.key.width = unit(3, units = "mm"),
        legend.background = element_blank())+
  ggspatial::annotation_scale(location = "tl")+
  coord_sf(crs = 32631)+
  labs(fill = "inferred\nduration (min)")
```

@@

```{r}
# get known patch summary
patch_summary_real <- data_res[, list(nfixes_real = .N,
                                      x_median = round(median(x), digits = -2),
                                      y_median = round(median(y), digits = -2),
                                      duration_real = as.numeric(
                                        max(time) - min(time))), 
                               by = "tID"]

@@

```{r}
# make figure comparing different methods
figure_patch_duration <-
  ggplot()+
  geom_segment(data = patch_summary_compare,
               aes(x = 100,
                   xend = duration,
                   y = duration_real,
                   yend = duration_real),
               col = "grey")+
  geom_segment(data = patch_summary_compare,
               aes(x = duration,
                   xend = duration,
                   y = 100,
                   yend = duration_real),
               col = "grey")+
  geom_point(data = patch_summary_compare,
             aes(duration, 
                 duration_real),
             show.legend = F,
            size = 3,
            stroke = 1,
            shape = 21,
            fill = pal[2],
            col = "white")+
  geom_smooth(data = patch_summary_compare,
             aes(duration, 
                 duration_real),
             se = F,
             size = 0.5,
             col = pal[1],
             alpha = 0.2,
             method = "glm",
             show.legend = F)+
  annotate(geom = "text",
           x = 300, y = 900,
           size = 5,
           label = "R^2 == 0.849",
           parse = T)+
  scale_x_log10(breaks = c(300, 600, 900),
                labels = as.integer(c(300, 600, 900) / 60))+
  scale_y_log10(breaks = c(300, 600, 900),
                labels = as.integer(c(300, 600, 900) / 60))+
  coord_equal(expand = F,
              xlim = c(200, 1200),
              ylim = c(200, 1200))+
  ggthemes::theme_few()+
  labs(x = "inferred duration (min)",
       y = "real duration (min)")
```

>>>>

# plot data
fig_data_raw <-
  ggplot(data)+
  geom_path(aes(x, y),
             col = "grey", alpha = 0.5, size = 0.3)+
  geom_point(aes(x, y),
             col = pal[3], alpha = 0.2, size = 2)+
  ggthemes::theme_map()+
  coord_sf(crs = 32631)

@@

```{r}
# divide by 1000, convert to integer, then convert to POSIXct
data[, time := as.integer(TIME / 1000)]
```

@@

```{r}
# plot filtered data
fig_speed_outliers <-
  ggplot()+
  geom_point(data = data,
             aes(x, y, 
                 col = speed_in < 25.6),
             size = 0.3, alpha = 0.5,
             show.legend = F)+
  scale_color_manual(values = c("grey", pal[3]))+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "grey99"))+
  ggspatial::annotation_scale(location = "tl")+
  coord_sf(crs = 32631)+
  labs(colour = "speed (m/s)")

# save
ggsave(fig_speed_outliers, filename = "figures/fig_speed_outlier.png",
       width = 170 / 25, height = 170 / 25)
```

@@

Here we use a speed threshold of 25.6 m/s, a little lower than the 95th percentile.

@@

# remove speed outliers
data <- atl_filter_covariates(data = data,
            filters = c("(speed_in < 15 & speed_out < 15)"))

@@

```{r}
# make a data copy
data_unproc <- copy(data)
```

@@

Subset for boating segments: 2020-08-24 11:13:00 -- 2020-08-24 12:36:11.

```{r}
# load recurse
library(recurse)
data_recurse <- data_unproc[, list(x, y, time, TAG, UTCtime)]

# filter for boating segment
data_recurse <- atl_filter_covariates(data = data_recurse,
  filters = c("between(UTCtime, '2020-08-24 11:13:00', '2020-08-24 12:36:11')"))

# get 4 column data
data_recurse <- data_recurse[, list(x, y, time, TAG)]

# get recurse data for a 10m radius
recurse_stats <- getRecursions(data_recurse,
                               radius = 500, timeunits = "mins")

# assign to recurse data
data_recurse[, res_time := recurse_stats$residenceTime]
```

### Plot residence time

```{r}
fig_calib_residence <- 
  ggplot()+
  geom_point(data = data_recurse,
             aes(x, y, col = res_time),
             shape = 19, alpha = 0.5,
             show.legend = F)+
  scale_colour_distiller(palette = "Greens", direction = 1,
                         trans = "log10",
                         labels = scales::comma,
                         limits = c(5, NA),
                         na.value = "grey")+
  ggspatial::annotation_scale(location = "br")+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "grey99"))+
  coord_sf(crs = 32631)
```

@@

```{r}
# assign id as tag
data_recurse[, id := as.character(TAG)]

# subset on 5 minute residence time
data_recurse_subset <- data_recurse[res_time >= 5, ]

# on known residence points
patch_res_known <- atl_res_patch(data_recurse_subset, 
                                buffer_radius = 5,
                                lim_spat_indep = 100,
                                lim_time_indep = 30,
                                min_fixes = 3)
```

@@

```{r}
# patch with residence points and all patches
fig_basic_residence <-
  ggplot()+
  geom_point(data = data_recurse,
            aes(x, y,
                col = res_time > 5),
            show.legend = F,
            size = 1)+
  geom_path(data = patch_summary_data,
            aes(x_median, y_median),
            col = pal[4], lty = 1)+
  geom_point(data = patch_summary_data,
             aes(x_median, y_median),
             size = 10,
             stroke = 2,
             col = pal[4],
             shape = 1,
             show.legend = F)+
  scale_colour_manual(values = c("grey", pal[3]))+
  scale_fill_distiller(palette = "RdPu")+
  ggspatial::annotation_scale(location = "tl")+
  ggthemes::theme_few()+
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "grey99"))+
  coord_sf(crs = 32631,
           xlim = c(NA, 650000))
```

@@

```{r}
# get known patch summary
data_res <- data_unproc[stringi::stri_detect(tID, regex = "(WP)"), ]
# filter for boating
data_res <- data_res[between(UTCtime, '2020-08-24 11:13:00', '2020-08-24 12:36:11'), ]

# get waypoint summary
patch_summary_real <- data_res[, list(nfixes_real = .N,
                                      x_median = round(median(x), digits = -2),
                                      y_median = round(median(y), digits = -2),
                                      duration_real = as.numeric(
                                        max(time) - min(time))), 
                               by = "tID"]

@@

```{r}
# make figure comparing different methods
# figure_patch_duration <-
  ggplot()+
  geom_point(data = patch_summary_compare,
             aes(duration, 
                 duration_real),
             show.legend = F,
            size = 3,
            stroke = 1,
            shape = 21,
            fill = pal[2],
            col = "white")+
  geom_smooth(data = patch_summary_compare,
             aes(duration, 
                 duration_real),
             se = F,
             size = 0.5,
             col = pal[1],
             alpha = 0.2,
             method = "glm",
             show.legend = F)+
  annotate(geom = "text",
           x = 500, y = 320,
           size = 5,
           label = "R^2 == 0.849",
           parse = T)+
  # scale_x_log10(breaks = c(300, 600, 900),
  #               labels = as.integer(c(300, 600, 900) / 60))+
  # scale_y_log10(breaks = c(300, 600, 900),
  #               labels = as.integer(c(300, 600, 900) / 60))+
  # coord_equal(expand = F)+
  ggthemes::theme_few()+
  labs(x = "inferred duration (min)",
       y = "real duration (min)")
```
